---
title: "Penguins data exploration"
author: "Wendy Phillips"
date: "2023-08-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## Introduction

The purpose of this .Rmd is for me to enjoy exploring data about penguins and also to demonstrate some R methods. 
```{r}
library(tidyverse)
library(scatterPlotMatrix)
library(palmerpenguins)
```

Get data from penguins library.
```{r}
data("penguins")
```

Take a peek at the data.
```{r}
str(penguins)
```
Hmmm... we might not want year to be an integer. It might be better as a factor, but for now we'll leave it that way.   

\   

Take a different peak at the data.
```{r}
summary(penguins)
```

From the above two outputs, we can see that sex has 11 NAs and two factor levels. The measurement columns each have 2 NA values. I'd like to drop the rows with measurement NAs but not with sex NAs. I'll first change the sex NAs to a new factor level "unknown". Then I'll be able to use drop.na to drop the measurement NAs.

```{r}
# First change sex back to a non-factor
penguins$sex <- as.character(penguins$sex)

# Change the sex NAs to unknown
penguins$sex[is.na(penguins$sex)] <- "unknown"

# Change sex back tor a factor
penguins$sex <- factor(penguins$sex)
```

Check if that did what I wanted.
```{r}
table(penguins$sex)
```

It did. Now I can use drop_na to get rid of the measurement NA rows.
```{r}
penguins <- drop_na(penguins)
```

Check the data again.
```{r}
summary(penguins)
```
For a very quick peek at relationships between numeric features, we can use the base R graphics functions pairs.
```{r}
pairs(select_if(penguins, is.numeric))
```
For a much more sophisticated visualization of the relationship between features, we can use scatterPlotMarix. Here is a plot with mostly default paramaters.
```{r}
scatterPlotMatrix(penguins, zAxisDim = "species", distribType = 1)
```

The column labels are difficult to see because they overlap so much. I'll change the "_" characters to a break to make the names fit better.
```{r}
columnLabels <- gsub("_", "<br>", colnames(penguins))
scatterPlotMatrix(penguins, zAxisDim = "species", distribType = 1, columnLabels = columnLabels)
```

Record versions being used.
```{r}
sessionInfo()
```

