---
title: "Colors in R"
author: "Wendy Phillips"
date: '2022-09-09'
output: html_document
---

# This is a work in progress!!!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(viridis)
library(RColorBrewer)
library(scales)
library(rcartocolor)
```


## Working with colors in R

Sometimes you want to explore what a palette of colors looks like. Here are some ways of plotting them.

First make a set of colors.
```{r}
colors <- c ("blue", "green", "purple", "red", "orange", "pink")
```

The library **scales** has a function that does this nicely.
```{r}
scales::show_col(colors)
```

You can show the colors in a pie chart, as shown below.
```{r}
pie(rep(1,6), col=colors)
```

You could create your own function.
```{r}
showCols <- function(cl=colors(), bg = "white", cex = 1.5, rot = 30) {
  m <- ceiling(sqrt(n <-length(cl)))
  length(cl) <- m*m; cm <- matrix(cl, m)
  require("grid")
  grid.newpage(); vp <- viewport(w = .5, h = .5)
  grid.rect(gp=gpar(fill=bg))
  grid.text(cm, x = col(cm)/m, y = rev(row(cm))/m, rot = rot,
            vp=vp, gp=gpar(cex = cex, col = cm))
}
```


```{r}
showCols(colors)
```

### Built in graphics device colors
The R graphics device has a built in set of colors. We can explore those.
```{r}
# Get all non-grey colors
non_grey_colors = grDevices::colors()[grep('gr(a|e)y', grDevices::colors(), invert = T)]
length(non_grey_colors)
```

```{r}
# Get all non-grey colors
color = grDevices::colors()[grep('gr(a|e)y', grDevices::colors(), invert = T)]

# Randomly sample from those colors
cols <- sample(color, 26)
scales::show_col(cols, cex_label = 0.5)
```

How many colors in the grDevices color palette include 'blue' in the name?
```{r}
# Get all non-grey colors
blues = grDevices::colors()[grep('blue', grDevices::colors())]
length(blues)
scales::show_col(blues, cex_label = 0.5)
```

```{r}
jet.colors <- colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan", "#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"))

scales::show_col(jet.colors(20), cex_label = 0.5)
```



The two most commonly used libraries with their own sets of color palettes are **RColorBrewer** and **viridis**.
Make color palettes.  

Some color libraries have built in functions for displaying their colors.

For RColorBrewer:
```{r}
RColorBrewer::display.brewer.pal(9, "Greens")
```


```{r}

heatmap_cols <- viridis(n = 256, alpha = 1, begin = 0, end = 1, 
                        direction = 1, option = "D")

```



```{r}

col=brewer.pal(n=12, name="Set3")

rgb.palette <- colorRampPalette(c("red", "green", "blue"),
                                space = "rgb")
mycolors <- colorRampPalette(brewer.pal(11,"BrBG"))

rain.palette <- colorRampPalette(brewer.pal(9,"Blues"))
pie(rep(1,6), col=rain.palette(6))

getPalette = colorRampPalette(brewer.pal(9, "Set1"))
pie(rep(1,10), col=getPalette(10))

```

```{r}
display.brewer.pal(8, "Set1")
```

```{r}
brewer.pal(9, "Greens")[c(4,6,8)]
```






```{r}
# From https://carto.com/carto-colors/
safe_CARTOColors <- c("#88CCEE","#CC6677","#DDCC77","#18ad4a","#332288","#AA4499","#44AA99","#999933","#882255","#661100","#6699CC")

pie(rep(1,11), col=safe_CARTOColors)
# Colors for metadata cats
bmi_group_cols <- c(safe_CARTOColors[[1]], safe_CARTOColors[[5]])
ft_mass_to_ht_gp_cols <- c(safe_CARTOColors[[2]], safe_CARTOColors[[10]])
ag_ratio_gp_cols <- c(safe_CARTOColors[[3]], safe_CARTOColors[[8]])


```